---
title: "published-code"
output: html_document
---

# REQUIRED PACKAGES AND FUNCTIONS

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
library(janitor)
library(lubridate)
library(gtools)
library(scales)
options(digits=6)
options(scipen=999)

'%notin%' <- Negate('%in%')


#does a standard group_by and count() with sum
grouper_sum <- function(input_df, group_by_column, sum_column, new_column_name = "n()"){
  output_df <- input_df %>%
    group_by(.data[[group_by_column]]) %>%
    summarise(temp_count = sum(.data[[sum_column]])) %>%
    mutate(percent = temp_count/sum(temp_count)*100) %>%
    arrange(desc(percent)) %>%
    rename(!!new_column_name := temp_count)
  return(output_df)
}

#does a standard group_by and count() with percentage
grouper <- function(input_df, group_by_column, new_column_name = "n()"){
  output_df <- input_df %>%
    group_by(.data[[group_by_column]]) %>%
    summarise(temp_count = n()) %>%
    mutate(percent = temp_count/sum(temp_count)*100) %>%
    arrange(desc(percent)) %>%
    rename(!!new_column_name := temp_count)
  return(output_df)
}

#group/counts every column in input dataframe
group_count <- function(input_df, group_column_name='n()', state_filter=NA, start_col = 1){
  column_names <- colnames(input_df)
  if(!is.na(state_filter)){
    input_df <- input_df %>%
      filter(state == state_filter)
  }
  for (column in column_names[start_col:length(column_names)]){
    output <- grouper(input_df, column, group_column_name)
    print(output)
  }
}

#function for calculating age, updates to the day; stolen from the internet
calc_age <- function(birthDate, refDate = Sys.Date()) {
    require(lubridate)
    period <- as.period(interval(birthDate, refDate),
                        unit = "year")
    period$year
}

#lowers case of every character column in a dataframe
lower_df <- function(input_df){
  names <- colnames(input_df)
  output_df <- input_df
  names <- colnames(output_df)
  for (name in names){
    if (is.character(output_df[[name]])){
      output_df[[name]] <- tolower(output_df[[name]])
      #print('yes')
    } else {
      output_df[[name]] <- output_df[[name]]
      #print('no')
    }
  }
  return(output_df)
}

#imports every file in a folder
import_files_in_folder <- function(input_path, output_name=''){
  files <- list.files(path=input_path)
  for (file in files){
    print(file)
    file_path = paste0(input_path, file)
    file_sans_csv = str_remove(file, '.csv')
    file_sans_csv = gsub("-", "_", file_sans_csv)
    imported_file <- read_csv(file_path)
    assign(paste0(file_sans_csv, output_name), imported_file, envir = parent.frame())
  }
}


pivot <- function(input_df, input_type){
  output_df <- input_df %>%
    select(-percent) %>%
    mutate(type = input_type) %>%
    relocate(type)
  return(output_df)
}

```

# REPAIR TABULA EXPORTS
```{python}
# The Purdue PDF was turned into 42 csvs using Tabula.
# Regex functions were used to clean most errors created by Tabula, that code is:

class Cleaner:
    
    def clean(self, input_data):
        #removes quotation marks at the front and end of numbers
        self.output_data = re.sub(r'(?!(([^"]*"){2})*[^"]*$),', '', input_data)
        #removes '.00's and replaces round brackets with negatives  
        self.output_data = self.output_data.replace('(', '-').replace(')', '').replace('.00', '').replace('"', '')
        #removes commas in numbers
        self.output_data = re.sub(r'(\w)\s([0-9]+.[0-9]+),', r'\1,\2 ', self.output_data)
        return self.output_data

for dirty_file in dirty_files:
    print(dirty_file)
    with open(dirty_file) as file:
        data = file.read()
        file.close()
    cleaner = Cleaner()
    clean_data = cleaner.clean(data)
    output_name = dirty_file.split('/')[-1].replace('  ', ' ').replace(' ', '-').replace('---', '-').replace('--', '-')
    with open(f'./data/cleaned-tabula/{output_name}', 'w') as output:
        output.write(clean_data)
        output.close()
```

# IMPORT AND CLEAN TABULA

```{r}

# Each csv was hand inspected and matched to the original PDF. 
# This cell imports and binds all of the cleaned tabula csv outputs.
# After binding, the entire sheet will be checked again for erros.

#this function was slightly modified to force it to import the list of cleaned tabula files in order so that they will be bound in order
import_bind_files_in_folder <- function(input_path, output_name=''){
  files <- mixedsort(sort(list.files(path=input_path)), decreasing = TRUE)
  master <- read_csv(paste0(input_path, files[1]))
  for (file in files[2:length(files)]){
    binder <- read_csv(paste0(input_path, file))
    master <- master %>%
      rbind(binder)
  }
  return(master)
}

#importing and binding
import_master <- import_bind_files_in_folder('./data/cleaned-tabula/') %>%
  #cleaning headers
  clean_names() %>%
  #making these columns numeric
  mutate(x2002 = as.numeric(x2002),
         x2003 = as.numeric(x2003),
         x2004 = as.numeric(x2004),
         x2013 = as.numeric(x2013),
         x2014 = as.numeric(x2014),
         )

#removing NAs
import_master[is.na(import_master)] <- 0

#writing to csv for another check and pivoting in Python
#write_csv(import_master, './data/binded_cleaned_tabula.csv')
```

#PIVOT TO DATABASE OF PAYMENTS

```{PYTHON}

from ..base.utils import Utilities
from tqdm import tqdm

class PaymentParser:
    
    # This parser was made to pivot Purdue Pharma payments from a list of payments in an accounting spreadsheet to a database of payments made to organizations.
    # This data was a PDF that was turned into a CSV using Tabula. The output of Tabula was corrected using regex expressions and hand modifications.
    # Generally, this PDF lists organization names in uppercase and departments/expense types in normal case. The pivot logic takes advantage of this format.
        # There are known errors in this logic. They have been compensated for in CAPS_DEPARTMENTS and LOWER_ORGS.

    #provides minor utilities
    utils = Utilities()
    #header of output database
    HEADER = ['organization', 'department', 'year', 'amount']
    #list of known departments that are in all caps
    CAPS_DEPARTMENTS = ['LELE', 'ISQA', 'NDSE', 'PKDM', 'HECON', 'PPMD', 'PPMD - QA', 'PPMD - IT']
    #list of known orgs that are not in all caps
    LOWER_ORGS = ['American College of Toxicology', 'AVENTINE CO dba AVENTINE HEALTH SCIENCES', 'Drug Enforcement Agents Recreation Association', 'Maine Chiefs of Police Association', 'National Fibromyalgia & Chronic Pain Assn', 'Police Chiefs of Rockland County', 'Pain Connection']

    # opens csv and breaks it into a list of lines
    def _open(self, input_file):
        self.input_file = input_file
        with open(self.input_file) as file:
            self.input_data = file.readlines()

    #sets the year of a payment based on column position (enumerable points in the for loop)
    def _set_year(self, nume):
        self.nume = nume
        if self.nume == 0:
            self.output_year = 2002
        elif self.nume == 1:
            self.output_year = 2003
        elif self.nume == 2:
            self.output_year = 2004
        elif self.nume == 3:
            self.output_year = 2005
        elif self.nume == 4:
            self.output_year = 2006
        elif self.nume == 5:
            self.output_year = 2007
        elif self.nume == 6:
            self.output_year = 2008
        elif self.nume == 7:
            self.output_year = 2009
        elif self.nume == 8:
            self.output_year = 2010
        elif self.nume == 9:
            self.output_year = 2011
        elif self.nume == 10:
            self.output_year = 2012
        elif self.nume == 11:
            self.output_year = 2013
        elif self.nume == 12:
            self.output_year = 2014
        elif self.nume == 13:
            self.output_year = 2015
        else:
            raise Exception('Enumerable is out of range.')
        return self.output_year

    #tests if string is in list of known departments that are in all caps
    def is_hecon(self, hecon_line):
        self.hecon_line = hecon_line
        if self.hecon_line in self.CAPS_DEPARTMENTS:
            return True
        else:
            return False
    
    #tests if string is UPPER or in list of known orgs that are not in all UPPER
    def is_upper(self, upper_line):
        self.upper_line = upper_line
        if self.upper_line.isupper() == True:
            return True
        elif self.upper_line in self.LOWER_ORGS:
            return True
        else:
            return False

    #assigns the organization name if the first column in the line/row passes the is_upper test 
    def _parse_upper(self, line):
        self.line = line
        self.org = self.utils.clean(self.line[0])

    #writes each column in a row to pivoted database, assigned to last org name assigned in parse_upper
    def _parse_lower(self, line):
        self.line = line
        self.dept_type = self.utils.clean(self.line[0])
        for self.nume, self.entry in enumerate(self.line[1:15]):
            self.output_year = self._set_year(self.nume)
            self.data_out = [self.org, self.dept_type, self.output_year, self.entry]
            self.utils.write_to_csv(self.data_out, self.output_csv)
    
    #entrypoint
    def parse_lines(self, input_file, output_file, output_file_path=''):
        #assigns inputs to self
        self.input_file = input_file
        self.output_file = output_file
        self.output_file_path = output_file_path
        #opens file
        self._open(self.input_file)
        #creates output csv
        self.output_csv = self.utils.create_csv(self.output_file, self.HEADER, self.output_file_path)
        #for each line in the database (tqdm provides a pretty progress bar)
        for self.line in tqdm(self.input_data[1:]):
            #split line into list
            self.line = self.line.split(',')
            #test list[0] to see if it is passes organization tests
            self.upper_test = self.is_upper(self.line[0])
            self.hecon_test = self.is_hecon(self.line[0])
            if self.upper_test is True and self.hecon_test is False:
                #if is an org, set the current self.org for the name
                self._parse_upper(self.line)
            else:
                #if it is not an org, log each payment for the last org and write it to the output csv
                self._parse_lower(self.line)

```

#CLASSIFYING ORGS

```{r}

#importing broad classifications that were made using openrefine and hand-sorting and labeling organizations
broad_classifications <- read_csv('./data/cleaning-steps/broad-analysis-hand-coded-refined.csv') %>%
  clean_names() %>%
  rename(position_other = medical_position_other) %>%
  select(organization, broad_org_type, state, position_other, medicine_type) %>%
  mutate(medicine_type = case_when(
    medicine_type == 'anesthesiology' ~ 'anesthesia',
    medicine_type == 'anethesia' ~ 'anesthesia',
    medicine_type == "oncolocy" ~ "oncology",
    medicine_type == "osteopathy" ~ "osteopathic",
    medicine_type == 'rehabilitation' ~ 'physical/rehab',
    organization == 'nc charlotte chapter of napnap' ~ 'pediatric',
    organization == 'new hampshire board of pharmacy' ~ 'general',
    medicine_type == 'cancer' ~ 'oncology',
    medicine_type == 'general' ~ 'other',
    medicine_type == 'narcotic' ~ 'narcotics',
    is.na(medicine_type) ~ 'other',
    TRUE ~ medicine_type),
    state = case_when(
      state == 'american' ~ 'none',
      is.na(state) ~ 'none',
      TRUE ~ state),
    position_other = case_when(
      is.na(position_other) ~ 'other',
      position_other == 'administratos' ~ 'administrator',
      position_other == 'academy of pai' ~ 'other',
      position_other == 'assn of drug diversi' ~ 'other',
      position_other == 'de la mujer latina in' ~ 'other',
      position_other == 'michigan chronic fat' ~ 'other',
      position_other == 'osteopathic medic' ~ 'other',
      position_other == 'patient access' ~ 'other',
      position_other == 'pharmacy' ~ 'pharmacist',
      position_other == 'research' ~ 'other',
      position_other == 'state board of pharmacy' ~ 'pharmacist',
      position_other == 'state hospice' ~ 'other',
      position_other == 'state medical' ~ 'other',
      position_other == 'attorneys general' ~ 'law enforcement',
      position_other == 'law enformcent' ~ 'law enforcement',
      position_other == 'law encofremen' ~ "law enforcement",
      position_other == 'nurses' ~ "nurse",
      position_other == 'nursing' ~ "nurse",
      organization == "democratic governors association" ~ "legislative",
      organization == 'national vulvodynia association' ~ '',
      TRUE ~ position_other)) %>%
  mutate(broad_org_type = case_when(
      broad_org_type == 'hospice & palliative care' ~ 'hospice and palliative care',
      broad_org_type == 'hospice and palliative care association' ~ 'hospice and palliative care',
      broad_org_type == 'hospice and palliative care organization' ~ 'hospice and palliative care',
      broad_org_type == 'gerontological society of america' ~ 'gerontological society',
      broad_org_type == 'gerontology society' ~ 'gerontological society',
      broad_org_type == 'lung assoc' ~ 'lung association',
      broad_org_type == 'osteopathic assn' ~ 'osteopathic association',
      broad_org_type == 'pharmacy association' ~ 'pharmacists association',
      broad_org_type == 'assoc of rehab nurses' ~ 'association of rehab nurses',
      TRUE ~ broad_org_type))


#importing pivoted, binded, cleaned tabula, filtering out 0 payments, removing double spaces, add broad orgs, creating medical/non-medical, exporting for last openrefine pass

#creating list of medical
positions <- c('pharmacist', 'physician', 'nurse', 'physician assistant', 'patient advocacy', 'medical director', 'case manager', 'medical publication', 'director', 'administration', "doctor's association", 'nurse, physician assistants')

old_master <- read_csv('./data/cleaning-steps/pivoted_binded_cleaned_tabula.csv') %>%
  filter(amount > 0) %>%
  mutate(organization = gsub('  ', ' ', organization)) %>%
  left_join(broad_classifications, by = "organization") %>%
  rename(broad_org = broad_org_type) %>%
  mutate(type = case_when(
    position_other %in% positions ~ 'medical',
    TRUE ~ position_other),
    position_other = case_when(
    position_other %in% positions ~ position_other,
    TRUE ~ '')) %>%
  rename(position = position_other) %>%
  relocate(organization, broad_org, state, type, position, medicine_type)

#exporting for last openrefine pass
#write_csv(master, './data/joined_pivoted_binded_cleaned_tabula.csv')

```

## IMPORT AND RECLASSIFY FOR FINAL MASTER

```{R}
#import to reclassify to simpler classification model
master <- read_csv("data/cleaning-steps/hand_fixed_refined_joined_pivoted_binded_cleaned_tabula.csv") %>%
  mutate(department = gsub('  ', ' ', department)) %>%
  mutate(organization = case_when(
    organization == 'national patient advocate foundatio' ~ 'patient advocate foundation',
    TRUE ~ organization)) %>%
  group_by(organization) %>%
  ungroup() %>%
  mutate(type = case_when(
    organization == 'pharmaceutical research & manufacturers of america' ~ 'phrma',
    position %in% c('pharmacist', 'physician', 'nurse', 'physician assistant', 'medical director') & organization != 'pharmaceutical research & manufacturers of america' & subtype != 'pain' & subtype != 'oncology' & subtype != 'hospice' & broad_org != 'board of pharmacy' ~ 'medical societies',
      type == 'medical' & organization != 'pharmaceutical research & manufacturers of america' & subtype != 'pain' & subtype != 'oncology' & subtype != 'hospice' ~ 'healthcare',
    type == 'law enforcement' ~ 'law enforcement',
    type == 'legislative' ~ 'political',
    type == 'drug abuse prevention' ~ 'drug abuse prevention',
    type == 'settlement' ~ 'settlement',
    subtype == 'pain' ~ 'pain advocacy',
    subtype == 'oncology' ~ 'cancer and hospice',
    subtype == 'hospice' ~ 'cancer and hospice',
    TRUE ~ 'other groups'
  )) %>%
  mutate(position = case_when(
    type == "medical societies" ~ position,
    TRUE ~ 'NA'
  )) %>%
  mutate(medicine_type = case_when(
    type == "medical societies" ~ subtype,
    TRUE ~ 'NA'
  )) %>%
  rename(medical_position = position) %>%
  select(organization, broad_org, state, type, medical_position, medicine_type, department, year, amount)

#export to csv to repair missing and incorrect data
#write_csv(master, './data/master-output-last-clean.csv')
```

## FINAL CLEANING PASS, IMPLIMENT FINAL CLASSIFCATION NAMES

```{R}

#24 organizations had incorrect or missing data
#importing corrected master
master <- read_csv('./data/master-output-last-clean-repaired.csv') %>%
  clean_names() %>%
  select(-x10) %>%
  rename(old_type = type)

#importing final corrected type
types <- read_csv('./data/purdue-pharma-corrected-type.csv') %>%
  select(-last_payment, -total)

master <- master %>%
  left_join(types, by = "organization") %>%
  select(organization, broad_org, state, type, medical_position, medicine_type, department, year, amount)

#writing outfile to assign subtypes
pain_groups <- master %>%
  filter(type == "pain advocacy") %>%
  group_by(organization, type) %>%
  summarise(total = sum(amount))
#write_csv(pain_groups, './data/pain-groups-type.csv')

#reading cleaned pain subtypes
pain_subtypes <- read_csv('./data/pain-groups-type-JL.csv') %>%
  select(-type, -total) %>%
  rename(pain_type = pain_subtype)

#final classifications chosen, renaming, joining pain subtypes
master <- master %>%
  mutate(type = case_when(
    type == "medical societies" ~ "other professional medical societies",
    type == "pain advocacy" ~ "pain groups",
    type == "healthcare" ~ "other healthcare",
    type == "settlement" ~ "wv settlement",
    TRUE ~ type
  )) %>%
  left_join(pain_subtypes) %>%
  select(organization, broad_org, state, type, medical_position, medicine_type, pain_type, department, year, amount) 

master <- master %>%
  mutate(broad_org = case_when(
    organization == "american cancer society" & broad_org == "conventions" ~ "cancer society",
    TRUE ~ broad_org
  ))
  

ons_changes<- master %>%
  filter(broad_org != "oncology nursing society" & str_detect(broad_org, "oncology nursing society"))

ons_changes <- ons_changes$broad_org

master <- master %>%
  mutate(broad_org = case_when(
    broad_org %in% ons_changes ~ "oncology nursing society",
    TRUE ~ broad_org
  ))

master <- master %>%
  mutate(organization = case_when(
    broad_org == "patient advocate foundation" & state == "national" ~ "national patient advocate foundation",
    TRUE ~ organization
  ))

print(paste0('This sum - ', sum(master$amount), ' - should equal the total from the PDF: 115,788,973.23.'))

#write_csv(master, './pub-output/purdue-pharma-master.csv')


```

# ANALYSIS AND FACT-CHECK

###	'The biggest individual recipient by far was the Pharmaceutical Research and Manufacturers of America. With a total of $23 million, the powerful lobbying group made up roughly a fifth of Purdue’s third-party spending. Prior to 2007, almost all of the spending was denoted as the “Public Affairs” department. But from 2008 to 2013, spending doubled and was now described as “Fed Gov Affairs.”'

```{r}

phrma <- master %>%
  filter(type == "phrma")

grouper_sum(phrma, 'department', 'amount', 'amount')

```

```{r}

ggplot(data=phrma, aes(x=factor(year), y=amount, fill=department)) + 
  geom_bar(stat="identity", alpha=.65, color="black") + 
  theme_light() +
  #ggtitle("Total Purdue dollars spent by type, 2002 to 2015") + 
  ylab("Spending in dollars") + 
  xlab("year") + 
  theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust=1)) +
  scale_y_continuous(label=comma)

```

### PAIN ADVOCACY GROUPS

```{r}

pain <- master %>%
  filter(str_detect(type,"pain"))

sum(pain$amount)

```

```{r}

pain_count <- pain %>%
  group_by(organization, pain_type) %>%
  count() %>%
  ungroup() 

grouper(pain_count, "pain_type")

```


```{r}

grouper_sum(pain, "pain_type", "amount", "amount")

```

```{r}

american_pain_foundation <- master %>%
  filter(organization == "american pain foundation")

sum(american_pain_foundation$amount)
```

```{r}

american_pain_society <- master %>%
  filter(organization == "american pain society")

sum(american_pain_society$amount)
```

```{r}

center_for_practical_bioethics <- master %>%
  filter(organization == "center for practical bioethics inc")

sum(center_for_practical_bioethics$amount)
```

### OTHER PROFESSIONAL MEDICAL SOCIETIES

```{R}

medical_societies <- master %>%
  filter(type == "other professional medical societies")

sum(medical_societies$amount)

```

```{r}

grouper_sum(medical_societies, "organization", "amount", "amount")

```

```{r}

ama <- master %>%
  filter(organization %in% c('american medical association'))

sum(ama$amount)

```

```{r}

american_osteopathic_association <- master %>%
  filter(organization == "american osteopathic assn")

sum(american_osteopathic_association$amount)
```

#NATIONAL ASSOC OF DRUG DIVERSION INVESTIGATORS

```{r}

naddi <- master %>%
  filter(organization == "naddi national assoc of drug diversion investigators in")

sum(naddi$amount)
```

```{r}

naddi_by_year <- grouper_sum(naddi, "year", "amount", "amount") %>%
  arrange(year)

for (year_ in naddi_by_year$year){
  temp <- naddi_by_year %>%
    filter(year == year_)
  print(paste0(year_, ' - ', temp$amount))
}
print(naddi_by_year$amount)
```

### AMERICAN ACADEMY OF FAMILY PHYSICIANS

```{r}

aafp <- master %>%
  filter(organization == "american academy of family physicians")

sum(aafp$amount)
```

```{r}

aafp_by_year <- grouper_sum(aafp, "year", "amount", "amount") %>%
  arrange(year)

for (year_ in aafp_by_year$year){
  temp <- aafp_by_year %>%
    filter(year == year_)
  print(paste0(year_, ' - ', temp$amount))
}
```

### U.S. CONFERENCE OF MAYORS

```{r}

mayors <- master %>%
  filter(organization == "united states conference of mayors mayors business council")

sum(mayors$amount)
```

```{r}

mayors_by_year <- grouper_sum(mayors, "year", "amount", "amount") %>%
  arrange(year)

for (year_ in mayors_by_year$year){
  temp <- mayors_by_year %>%
    filter(year == year_)
  print(paste0(year_, ' - ', temp$amount))
}
```

### AMERICAN CANCER SOCIETY

```{r}

aca <- master %>%
  filter(broad_org == "cancer society")

grouped_aca <- aca %>%
  group_by(organization) %>%
  count()

aca_orgs <- grouped_aca$organization

aca_only <- aca %>%
  filter(organization == "american cancer society")

aca_network <- aca %>%
  filter(organization == "american cancer soc cancer ntwk inc")


aca_alliance <- aca %>%
  filter(organization == "american cancer society am alliance of cancer pain init")

aca_armstrong <- aca %>%
  filter(organization == "american cancer society armstrong unit")

sum(aca$amount)
```

```{r}

aca_only_by_year <- grouper_sum(aca_only, "year", "amount", "amount") %>%
  arrange(year)

for (year_ in aca_only_by_year$year){
  temp <- aca_only_by_year %>%
    filter(year == year_)
  print(paste0(year_, ' - ', temp$amount))
}
```

```{r}

aca_network_by_year <- grouper_sum(aca_network, "year", "amount", "amount") %>%
  arrange(year)

for (year_ in aca_network_by_year$year){
  temp <- aca_network_by_year %>%
    filter(year == year_)
  print(paste0(year_, ' - ', temp$amount))
}
```

```{r}

aca_alliance_by_year <- grouper_sum(aca_alliance, "year", "amount", "amount") %>%
  arrange(year)

for (year_ in aca_alliance_by_year$year){
  temp <- aca_alliance_by_year %>%
    filter(year == year_)
  print(paste0(year_, ' - ', temp$amount))
}
```

```{r}

aca_armstrong_by_year <- grouper_sum(aca_armstrong, "year", "amount", "amount") %>%
  arrange(year)

for (year_ in aca_armstrong_by_year$year){
  temp <- aca_armstrong_by_year %>%
    filter(year == year_)
  print(paste0(year_, ' - ', temp$amount))
}
```


### ONCOLOGY NURSING SOCIETY 

```{r}

ons <- master %>%
  filter(broad_org == "oncology nursing society")

ons_only <- master %>%
  filter(organization == "oncology nursing society")

ons_foundation <- master %>%
  filter(organization == "ons foundation")


sum(ons$amount)
```

```{r}

ons_only_by_year <- grouper_sum(ons_only, "year", "amount", "amount") %>%
  arrange(year)

for (year_ in ons_only_by_year$year){
  temp <- ons_only_by_year %>%
    filter(year == year_)
  print(paste0(year_, ' - ', temp$amount))
}
```

### ONCOLOGY NURSING SOCIETY FOUNDATION

```{r}

ons_foundation_by_year <- grouper_sum(ons_foundation, "year", "amount", "amount") %>%
  arrange(year)

for (year_ in ons_foundation_by_year$year){
  temp <- ons_foundation_by_year %>%
    filter(year == year_)
  print(paste0(year_, ' - ', temp$amount))
}
```

### JOHNS HOPKINS UNIVERSITY

```{r}

johns_hopkins <- master %>%
  filter(broad_org == "johns hopkins university")

johns_hopkins_by_year <- grouper_sum(johns_hopkins, "year", "amount", "amount") %>%
  arrange(year)

for (year_ in johns_hopkins_by_year$year){
  temp <- johns_hopkins_by_year %>%
    filter(year == year_)
  print(paste0(year_, ' - ', temp$amount))
}
```

### Biotechnology Industry Organization

```{r}

bio <- master %>%
  filter(broad_org == "biotechnology industry organ")

bio_by_year <- grouper_sum(bio, "year", "amount", "amount") %>%
  arrange(year)

for (year_ in bio_by_year$year){
  temp <- bio_by_year %>%
    filter(year == year_)
  print(paste0(year_, ' - ', temp$amount))
}
```

```{r}

sum(bio$amount)

```

### LAW ENFORCEMENT

```{r}

law_enforcement <- master %>%
  filter(type == "law enforcement")

sum(law_enforcement$amount)
```

```{R}

grouper_sum(law_enforcement, 'organization', 'amount', 'amount')

```

### DRUG ABUSE

```{r}

drug_abuse <- master %>%
  filter(type == "drug abuse prevention")

```

```{r}

ggplot(data=drug_abuse, aes(x=factor(year), y=amount, fill=department)) + 
  geom_bar(stat="identity", alpha=.65, color="black") + 
  theme_light() +
  #ggtitle("Total Purdue dollars spent by type, 2002 to 2015") + 
  ylab("Spending in dollars") + 
  xlab("year") + 
  theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust=1)) +
  scale_y_continuous(label=comma)

```

## RECIPIENTS WITH KEY FUNDING CHANGES OVER TIME

```{r}

key_changes <- c('American Pharmacists Association', 'Partnership For A Drug Free America Inc', 'Patient Advocate Foundation', "American Osteopathic Assn", "Pharmaceutical Research & Manufacturers of America", "American Academy of Pain Management", "American Academy of Pain Medicine", "Center For Practical Bioethics Inc", "American Medical Association")

group_org <- function(key_change){
  temp <- pretty_version %>%
    filter(organization == !!key_change) 
  
  temp <- grouper_sum(temp, 'year', 'amount', 'total') %>%
    #mutate(organization = paste0(!!key_change)) %>%
    #relocate(organization) %>%
    pivot(key_change)
}


spending_by_key_changes <- group_org(key_changes[1])

for (key_change in key_changes[2:length(key_changes)]){
  
  joiner <- group_org(key_change)
  
  spending_by_key_changes <- spending_by_key_changes %>%
    bind_rows(joiner)
}

spending_by_key_changes <- spending_by_key_changes %>%
  pivot_wider(names_from = type, values_from = total) %>%
  arrange(year)

#write_csv(spending_by_key_changes, './pub-output/spending-by-key-changes.csv')

```


# PUBLICATION OUTPUTS

```{r}
grouped_master <- master %>%
  group_by(organization, type, medical_position, medicine_type) %>%
  summarise(last_payment = max(year),
            total = sum(amount))


grouped_medical <- master %>%
  group_by(type) %>%
  summarise(total = sum(amount))


#COMMENTIING OUT CODE THAT WRITES PUB FILES

#write_csv(grouped_master, './pub-output/purdue-pharma-grouped-master.csv')
#write_csv(grouped_medical, './pub-output/purdue-pharma-grouped-medical.csv')
    
```

## PRETTY VERSION FOR FLOURISH HIERARCHY VISUALIZATION

```{r}
simpleCap <- function(x) {
  s <- strsplit(x, " ")[[1]]
  paste(toupper(substring(s, 1,1)), substring(s, 2),
      sep="", collapse=" ")
}



pretty_version <- master %>%
  select(-broad_org) %>%
  mutate(type = case_when(
    type == "wv settlement" ~ "WV Settlement",
    type == "phrma" ~ "PhRMA",
    TRUE ~ type)) %>%
  mutate(organization = sapply(organization, simpleCap),
         state = sapply(state, simpleCap),
         type = sapply(type, simpleCap),
         medical_position = sapply(medical_position, simpleCap),
         medicine_type = sapply(medicine_type, simpleCap),
         pain_type = sapply(pain_type, simpleCap)
         ) %>%
  mutate(organization = gsub("Of ", "of ", organization)) %>%
  mutate(organization = gsub("Assoc ", "Association", organization)) %>%
  mutate(organization = gsub("Soc ", "Society", organization)) %>%
  #mutate(department = gsub("Ff ", "FF", department)) %>%
  mutate(medical_position = gsub("NANA", "", medical_position),
         medicine_type = gsub("NANA", "", medicine_type),
         pain_type = gsub("NANA", "", pain_type),
         state = gsub("", "", state)) %>%
  mutate(state = gsub('National', '', state)) %>%
  mutate(state = gsub('None', '', state)) %>%
  mutate(organization = gsub('Ymca', 'YMCA', organization)) %>%
  mutate(type = gsub('Ymca', 'YMCA', type)) %>%
  mutate(organization = trimws(organization, c("both")),
         state = trimws(state, c("both")),
         type = trimws(type, c("both")),
         medical_position = trimws(medical_position, c("both")),
         medicine_type = trimws(medicine_type, c("both")),
         department = trimws(department, c("both")),
         #last_payment = as.character(last_payment)
         ) 
  

pretty_grouped <- pretty_version %>%
  group_by(organization, type) %>%
  count() %>%
  select(-n)

#write_csv(pretty_version, './pub-output/pretty-version.csv')

#write_csv(pretty_grouped, './pub-output/pretty-grouped.csv')

```

## FLOURISH BAR CHART TYPE VIZ

```{r}

type <- pretty_version %>%
  group_by(type) %>%
  summarise()

types <- type$type

spending_per_year <- grouper_sum(master, 'year', "amount", "dollars") %>%
  pivot('total')

spending_by_type <- spending_per_year %>%
  filter(type == 'none')

for (type_ in types){
  temp <- pretty_version %>%
    filter(type == !!type_) 
  joiner <- grouper_sum(temp, 'year', 'amount', 'dollars') %>%
    pivot(type_)
  spending_by_type <- spending_by_type %>%
    bind_rows(joiner)
}

spending_by_type <- spending_by_type %>%
  clean_names() %>%
  rename(total = dollars)

spending_by_type[is.na(spending_by_type)] <- 0 

ggplot(data=spending_by_type, aes(x=factor(year), y=total, fill=type)) + 
  geom_bar(stat="identity", alpha=.65, color="black") + 
  #scale_fill_manual(values = c("black", "red", "lightgreen", 'gold', 'blue', 'darkred', 'darkgreen')) +
  theme_light() +
  #geom_bar(aes(y=type), stat="identity", position ="identity", al pha=.3, fill='lightblue', color='lightblue4') +
  #geom_bar(aes(y=law_enforcement), stat="identity", position ="identity", alpha=.3, fill='red', color='red') +
  #geom_bar(aes(y=pharmacist), stat="identity", position ="identity", alpha=.3, fill='yellow', color='yellow') +
  #geom_bar(aes(y=physician), stat="identity", position ="identity", alpha=.3, fill='green', color='green') +
  ggtitle("Total Purdue dollars spent by type, 2002 to 2015") + 
  ylab("Spending in dollars") + 
  xlab("year") + 
  theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust=1)) +
  scale_y_continuous(label=comma)


```

```{r}

pretty_spending_by_type <- spending_by_type %>%
  pivot_wider(names_from = type, values_from = total) %>%
  filter(year != 2015) 


#write_csv(pretty_spending_by_type, './pub-output/spending-by-type.csv')

```

## FLOURISH BAR CHART DEPARTMENT VIZ

```{r}

spending_per_year <- grouper_sum(master, 'year', "amount", "dollars") %>%
  pivot('total')


department <- grouper_sum(pretty_version, "department", "amount", "dollars") %>%
  slice(1:30)

departments <- department$department

spending_by_department <- department %>%
  clean_names() %>%
  rename(total = dollars)

for (department_ in departments){
  spending_by_department_filtered <- pretty_version %>%
    filter(department == !!department_) 
  
  spending_by_department_filtered <- grouper_sum(spending_by_department_filtered, 'year', 'amount', 'total') %>%
    mutate(type = !!department_) %>%
    relocate(type)
  
  plot <- ggplot(data=spending_by_department_filtered, aes(x=factor(year), y=total)) + 
    geom_bar(stat="identity", alpha=.65, color="black") + 
    #scale_fill_manual(values = c("black", "red", "lightgreen", 'gold', 'blue', 'darkred', 'darkgreen')) +
    theme_light() +
    #geom_bar(aes(y=type), stat="identity", position ="identity", al pha=.3, fill='lightblue', color='lightblue4') +
    #geom_bar(aes(y=law_enforcement), stat="identity", position ="identity", alpha=.3, fill='red', color='red') +
    #geom_bar(aes(y=pharmacist), stat="identity", position ="identity", alpha=.3, fill='yellow', color='yellow') +
    #geom_bar(aes(y=physician), stat="identity", position ="identity", alpha=.3, fill='green', color='green') +
    ggtitle("Total Purdue dollars spent by type, 2002 to 2015") + 
    ylab("Spending in dollars") + 
    xlab("year") + 
    theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust=1)) +
    scale_y_continuous(label=comma)
  
  
  print(plot)

}

```
