---
title: "data-cleaning"
author: "Ryan Little"
date: "12/3/2019"
output: html_document
---

## R Markdown
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE, paged.print=TRUE)
```


## Load libraries

```{r}
library(tidyverse)  # attaches purrr and readr
library(lubridate)
library(rvest)
library(downloader)
library(R.utils)
library(rlist)
library(here)
library(janitor)
library(scales)
library(ggrepel)
library(ggplot2)
library(here)
library(tidycensus)
library(ggthemes)
library(scales)
library(mapview)
library(here)
library(svMisc)
library(blscrapeR)

# Store census API key
census_api_key("")

# the function that should be in Tidyverse but isn't
'%!in%' <- function(x,y)!('%in%'(x,y))

options(scipen=999)
options(digits=2)
```

```{r}

#clear environment for when I need it
rm(list = ls())

#why this function isn't in tidyverse, I do not know
'%!in%' <- function(x,y)!('%in%'(x,y))

#function for calculating age, updates to the day; stolen from the internet
calc_age <- function(birthDate, refDate = Sys.Date()) {
    require(lubridate)
    period <- as.period(interval(birthDate, refDate),
                        unit = "year")
    period$year
}

```


#Import and Clean Data

The data had to be downloaded by month by request of Okaloosa County Clerk. They said larger timeframes would tax their system. This means the data had to be downloaded by month, imported and binded together using a for loop.

###Missing Data

The underlying data this code works with was not included because of the personal information included in a database that includes all criminal charges from Okaloosa County, Florida from 2011 until June 2020. If you would like access to the underlying data, please email me at (ryanerinlittle@gmail.com)[mailto:ryanerinlittle@gmail.com].

```{r}

# list of file names
files <- c(
    "0110", "0210", "0310", "0410", "0510", "0610", "0710", "0810", "0910", "1010", "1110", "1210",
    "0111", "0211", "0311", "0411", "0511", "0611", "0711", "0811", "0911", "1011", "1111", "1211",
    "0112", "0212", "0312", "0412", "0512", "0612", "0712", "0812", "0912", "1012", "1112", "1212",
    "0113", "0213", "0313", "0413", "0513", "0613", "0713", "0813", "0913", "1013", "1113", "1213",
    "0114", "0214", "0314", "0414", "0514", "0614", "0714", "0814", "0914", "1014", "1114", "1214",
    "0115", "0215", "0315", "0415", "0515", "0615", "0715", "0815", "0915", "1015", "1115", "1215",
    "0116", "0216", "0316", "0416", "0516", "0616", "0716", "0816", "0916", "1016", "1116", "1216", 
    "0117", "0217", "0317", "0417", "0517", "0617", "0717", "0817", "0917", "1017", "1117", "1217", 
    "0118", "0218", "0318", "0418", "0518", "0618", "0718", "0818", "0918", "1018", "1118", "1218", 
    "0119", "0219", "0319", "0419", "0519", "0619", "0719", "0819", "0919", "1019", "1119", "1219",
    "0120", "0220", "0320", "0420", "0520", "0620")

# creating empty dataframe, changes to character were made playing wack-a-mole with binding errors
master_data <- read_csv(here("data", "0115.csv")) %>%
  clean_names() %>%
  filter(arrest_charge == "nothing") %>%
  mutate(actual_speed = as.character(actual_speed),
         blood_alcohol_level = as.character(blood_alcohol_level),
         commercial_vehicle = as.character(commercial_vehicle),
         haul_hazard = as.character(haul_hazard),
         obts_number = as.character(obts_number),
         vehicle_tag_expiration_year = as.character(vehicle_tag_expiration_year),
         prosecutor_charge_count = as.character(prosecutor_charge_count))

# importing, binding to empty data frame 
for (i in files) {
  #writing filepath
  name <- paste0(i, ".csv")
  filepath <- "data/"
  name_plus_filepath <- paste0(filepath,name)
  # importing and cleaning files to temp dataframe
  temp <- read_csv(name_plus_filepath) %>%
    clean_names() 
  temp <- unique(temp)
  # changes to character were made playing wack-a-mole with binding errors
  temp <- temp %>%
  mutate(actual_speed = as.character(actual_speed),
         blood_alcohol_level = as.character(blood_alcohol_level),
         commercial_vehicle = as.character(commercial_vehicle),
         haul_hazard = as.character(haul_hazard),
         obts_number = as.character(obts_number),
         vehicle_tag_expiration_year = as.character(vehicle_tag_expiration_year),
         prosecutor_charge_count = as.character(prosecutor_charge_count),
         zip = as.character(zip),
         bond_hearing_date = as.character(bond_hearing_date),
         arrest_charge_count = as.numeric(arrest_charge_count))
  # binding temp to master_data
  master_data <- master_data %>%
    bind_rows(temp)
  # processing and clearing environment
  rm(temp, name, filepath, name_plus_filepath)
}

rm(i)
rm(files)

# changing columns back to numerics, go lowercase
master_data <- master_data %>%
    mutate_all(tolower) %>%  
    mutate(actual_speed = as.double(actual_speed),
         blood_alcohol_level = as.double(blood_alcohol_level),
         commercial_vehicle = as.double(commercial_vehicle),
         haul_hazard = as.double(haul_hazard),
         obts_number = as.double(obts_number),
         vehicle_tag_expiration_year = as.double(vehicle_tag_expiration_year),
         prosecutor_charge_count = as.double(prosecutor_charge_count),
         case_open_date = mdy(case_open_date),
         case_close_date = mdy(case_close_date),
         disposition_date = mdy(disposition_date),
         date_of_birth = mdy(date_of_birth),
         offense_date = mdy(offense_date),
         arrest_date = mdy(arrest_date),
         arrest_month = month(arrest_date),
         arrest_year = year(arrest_date),
         offense_month = month(offense_date),
         offense_year = year(offense_date),
         close_year = year(case_close_date),
         total_assessed = as.numeric(total_assessed),
         balance = as.numeric(balance),
         total_paid = as.numeric(total_paid),
         obts_number = as.character(obts_number),
         arrest_charge_count = as.numeric(arrest_charge_count),
         arrest_week = week(arrest_date),
         next_day = (arrest_date + 1)
         ) %>%
    filter(offense_year %in% c("2010", "2011", "2012", "2013", "2014", "2015", "2016", "2017", "2018", "2019", "2020")) %>%
    select(-case_type, -disposition_date, -disposition, -defendant_code, -bond_hearing_date, -posted_speed, -actual_speed, -school_status, -commercial_vehicle, -haul_hazard, -d6_issued_date, -d6_issue_status, -d6_released_date)

#removing doubles that were created by the court's data download system
master_data <- master_data %>%
    unique()

#creating column with month case was opened
master_data <- master_data %>%
  mutate(case_open_month = substr(case_open_date, start = 1, stop = 7))

charges_by_month <- master_data %>%
  group_by(case_open_month) %>%
  summarise(count = n())

rm(temp)

# There are a few people who have errors in the data that have been manually cleaned:

# In every case, typos in name, birthday or drivers license were tested using a combination of the other two data points to prove the third

# Example: john doe (real name changed to protect personal info) had two different birthdays ("1977-12-XX" "1977-02-XX") were confirmed errors by # comparing driver's license; a mispelling of his name exists

master_data <- master_data %>%
  mutate(
    date_of_birth = case_when(
      defendant == "john doe" ~ as_date("1977-02-XX"),
      TRUE  ~ date_of_birth
    )
  ) %>%
  mutate(
    defendant = case_when(
      defendant == "john doe" ~ "john doe",
      TRUE  ~ defendant
    )
  ) %>%
  
# This example code was used to repair the errors in the database aproximately 40 times. That code has been deleted in this example to protect the personal information.
  

#importing scraped data from the Okaloosa County Jail website
jails1 <- read_csv("data/jail-stays-first.csv")
jails2 <- read_csv("data/jail-stays-second.csv")

#binding and removing entries with no name
jail_stays <-  jails1 %>%
  rbind(jails2) %>%
  filter(last != "none") %>%
  mutate(booking_date = mdy(booking_date),
         release_date = mdy(release_date),
         released_in_am = am(release_time),
         days_in_jail = replace_na((release_date - booking_date), 0),
         name = paste0(first, " ", last),
         arrest_week = week(booking_date),
         arrest_year = year(booking_date)
         )

rm(jails1, jails2)

```


```{r}
#importing and comparing state database aquired by the Howard Center
state_data <- read_csv("data/okaloosa-state-tres.csv")

state_data$court_docket_no <- tolower(state_data$court_docket_no)

anti_tres <- trespass_data %>%
  anti_join(state_data, by = c("uniform_case_number" = "court_docket_no"))

```

```{r}

other_anti <- state_data %>%
  anti_join(trespass_data, by = c("court_docket_no" = "uniform_case_number"))


```


# Filtering for Homeless
Fort Walton Beach Police officer confirms that "General Delivery" is an address law enforcement agencies use when they arrest a homeless person. Fort Walton Beach uses "At Large."

```{r}

# creating a dataframe that only includes addresses that match those we know signify they are homeless based on reporting
general_delivery <- master_data %>%
  filter(address %in% c("general delivery", "genral delivery", "at large", "genreal delivery", "general deliverly", "homeless", "gerneral delivery", "general deliver", "general delviery", "not listed", "general deivery", "grneral delivery", "428 mclaughlin ave", "307 harbor blvd", "183 eglin parkway ne", "117 windham ave", "300 miracle strip", "21 walter martin rd", "123 anywhere st"))


#filter for list of defendants
homeless_names <- general_delivery %>%
  group_by(defendant) %>%
  summarise(charges = n()) %>%
  arrange(desc(charges))

# converting names to vector to feed into for loop
names <- as.vector(homeless_names$defendant)

#creating blank dataframe to bind in the loop
homeless_data <- master_data %>%
  filter(defendant == "nothing")

for (i in names) {
  # filtering for person
  temp <- master_data %>%
    filter(defendant == i)
  # binding to dataframe
  homeless_data <- homeless_data %>%
    bind_rows(temp)
}

rm(temp, i, general_delivery, homeless_names)
```

# Trespass

This frame creates a dataframe that includes all trespass charges filed against all homeless people.

```{r}

# creating a dataframe just on trespass
# creating a list of all trespass charges 
all_charges <- master_data %>%
  group_by(arrest_charge, arrest_statute) %>%
  summarise(count = n())
# filtering for all charges with "tres" in the statute
tres_charges <- all_charges %>%
  filter(str_detect(arrest_statute, "tres"))
tres_charges <- tres_charges %>%
  select(arrest_charge) %>%
  unique()
# converting to vector to feed into for loop
trespass <- as.vector(tres_charges$arrest_charge)

# creating blank dataframe
trespass_data <- homeless_data %>%
  filter(arrest_charge == "nothing")

for (i in trespass) {
  # filter for trespass statute
  temp <- homeless_data %>%
    filter(arrest_charge == i)
  # bind to dataframe
  trespass_data <- trespass_data %>%
    bind_rows(temp)
}

trespass_by_month <- trespass_data %>%
  mutate(case_open_sort = substr(case_open_date, start = 1, stop = 7)) %>%
  group_by(case_open_sort) %>%
  summarise(count = n()) %>%
  inner_join(charges_by_month, by = c("case_open_sort" = "case_open_month")) %>%
  mutate(trespass_charges = count.x,
         total_charges = count.y) %>%
  select(-count.x, -count.y)

april_may <- trespass_data %>%
  mutate(arrest_sort = substr(arrest_date, start = 1, stop = 7)) %>%
  filter(arrest_sort %in% c("2020-04", "2020-05"))

# fWB police only

fwb_charges_by_month <- master_data %>%
  filter(arresting_agency == "ft walton bch police") %>%
  mutate(case_open_sort = substr(case_open_date, start = 1, stop = 7)) %>%
  group_by(case_open_sort) %>%
  summarise(charges = n())

fwb_charges <- master_data %>%
  #mutate(arrest_sort = substr(arrest_date, start = 1, stop = 7)) %>%
  #filter(arrest_sort %in% c("2020-03", "2020-04", "2020-05")) %>%
  group_by(arresting_agency) %>%
  count()




```




```{r}

# creating a dataframe just on trespass
# creating a list of all trespass charges 
all_charges <- master_data %>%
  group_by(arrest_charge, arrest_statute) %>%
  summarise(count = n())
# filtering for all charges with "tres" in the statute
tres_charges <- all_charges %>%
  filter(str_detect(arrest_statute, "tres"))
tres_charges <- tres_charges %>%
  select(arrest_charge) %>%
  unique()
# converting to vector to feed into for loop
trespass <- as.vector(tres_charges$arrest_charge)

# creating blank dataframe
test <- master_data %>%
  filter(arrest_charge == "nothing")


for (i in trespass) {
  # filter for trespass statute
  temp <- master_data %>%
    filter(arrest_charge == i)
  # bind to dataframe
  test <- test %>%
    bind_rows(temp)
}


test %>%
  group_by(prosecutor) %>%
  count() %>%
  arrange(desc(n))

```

```{r}
anti_test <- test_two %>%
  anti_join(test_three, by = c('uniform_case_number' = 'court_docket_no'))


```


```{r}

test_three <- state_data %>%
  filter(year >= '2010')



```


# Other Crimes

Through analysis and reporting, it became clear that other_crimes_list is a list of other state statutes that are used to criminalize homelessness and that a lot of people who are charged with trespassing also miss court a lot and get charged with resisting arrest. This code block creates a dataframe for each one.

```{r}
# creating dataframe for all other criminalization statutes that are not trespassing

# "337.406: improper use of state right of way

other_crimes_list <- c("316.130", "316.2045", "337.406", "856.011", "856.021", "877.03")


no_show_resist <- c( "843.15(1a)", "843.15(1a)*", "843.15(1b)", "843.15(1b)*", "843.02")

# 843.15(1a): bail secured bond-failure to appear for felony offense
# 843.15(1b): failure to appear for misdemeanor
# 843.15(1b)*: bail secured bond-failure to appear for misdemeanor offense
# 843.02: resist officer-obstruct wo violence"

other_crimes_data <- master_data %>%
  filter(arrest_charge == "nothing")

for (i in other_crimes_list) {
  temp <- homeless_data %>%
    filter(arrest_charge == i)
  print(paste0("Finished filtering ", i))
  other_crimes_data <- other_crimes_data %>%
    bind_rows(temp)
}

no_show_resist_data <- master_data %>%
  filter(arrest_charge == "nothing")

for (i in no_show_resist) {
  temp <- homeless_data %>%
    filter(arrest_charge == i)
  no_show_resist_data <- no_show_resist_data %>%
    bind_rows(temp)
}

rm(temp, i)

# this searches for the one state law found unconstitutional
unconst <- master_data %>%
  filter(str_detect(arrest_charge, "316.2045"))

# creating a spreadsheet of all the charges filed against homeless people
homeless_charges <- homeless_data %>%
  group_by(arrest_charge, arrest_statute) %>%
  summarise(count = n()) %>%
  arrange(desc(count))

# creating dataframe of trespass, other and resist_no_show crimes 
filtered_crimes <- trespass_data %>%
  rbind(other_crimes_data) %>%
  rbind(no_show_resist_data)

```


# Examing what the homeless are charged with most, building total impacts frame

## Demographics - complete

```{r}
#pulling demographic info 
temp <- homeless_data %>%
  group_by(defendant, race, gender, date_of_birth) %>%
  summarise()

demo_by_race <- temp %>%
  group_by(race) %>%
  summarise(count = n())
  #b 44 - 14%
  #w 267 - 85%
  #na 2 - 0.01 %
  #total 313
demo_by_gender <- temp %>%
  group_by(gender) %>%
  summarise(count =n())
  #43 women - 14%
  #270 men 86%
demographics <- temp %>%
  mutate(age = calc_age(date_of_birth)) %>%
  select(-date_of_birth)
 

```

```{r}

temp <- trespass_data %>%
  group_by(defendant, race, gender, date_of_birth) %>%
  summarise()

tres_demo_by_race <- temp %>%
  group_by(race) %>%
  summarise(count = n())



```

## Trespass Charges

```{r}
# calculating trespass fees
temp1 <- trespass_data %>%
  #select and unique to to clear cases with multiple charges that list the fees assessed for the entire case multiple times
  select(case_number, defendant, total_assessed) %>%
  unique() %>%
  group_by(defendant) %>%
  summarise(tres_fees = sum(total_assessed))

# calculating trespass charges
temp2 <- trespass_data %>%
  group_by(defendant) %>%
  summarise(tres_charges = n())

# calculating trespass fees paid
temp3 <- trespass_data %>%
  #select and unique to to clear cases with multiple charges that list the fees assessed for the entire case multiple times
  select(case_number, defendant, total_paid) %>%
  group_by(defendant) %>%
  summarise(tres_paid = sum(total_paid))

#create column that counts missed court charges that resulted from tresspassing
temp19 <- trespass_data %>%
  group_by(case_number, defendant) %>%
  summarise()

temp20 <- homeless_data %>%
  filter(str_detect(arrest_charge, "843.15")) %>%
  group_by(case_number) %>%
  summarise()

temp21 <- temp19 %>%
  inner_join(temp20, by = "case_number") %>%
  mutate(missed_court_charge = "yes")

tres_missed_court <- temp21 %>%
  group_by(defendant) %>%
  summarise(tres_missed_court = n())

# creating trespass impacts master frame
trespass_impacts <- temp1 %>%
  left_join(temp2) %>%
  left_join(temp3) %>%
  left_join(tres_missed_court) %>%
  select(defendant, tres_charges,tres_missed_court, tres_fees, tres_paid)
trespass_impacts$tres_missed_court[is.na(trespass_impacts$tres_missed_court)] <- 0

rm(temp, temp1, temp2, temp3) 

sum(trespass_impacts$tres_fees)

```


```{r}

# homeless non-trespass criminilzation charges
temp <- other_crimes_data  %>%
  select(case_number, defendant) %>%
  unique() %>%
  group_by(defendant, case_number) %>%
  summarise()
other_crimes <- temp %>%
  group_by(defendant) %>%
  summarise(other_crimes = n())
other_crimes_charges <- other_crimes_data  %>%
  select(case_number, defendant, arrest_charge_count) %>%
  unique() %>%
  group_by(defendant) %>%
  summarise(other_crimes_charges = n())

# summarizing court costs for other crimes
other_crim_fines <- other_crimes_data  %>%
  select(defendant, case_number, total_assessed, total_paid) %>%
  group_by(defendant) %>%
  summarise(other_crim_fines = sum(total_assessed))
total_fines <- homeless_data %>%
  select(defendant, case_number, total_assessed, total_paid) %>%
  group_by(defendant) %>%
  summarise(total_fines = sum(total_assessed))

# summarizing fees paid for other crimes
other_crim_paid <-other_crimes_data  %>%
  select(defendant, case_number, total_assessed, total_paid) %>%
  group_by(defendant) %>%
  summarise(other_crim_paid = sum(total_paid))
total_paid <- homeless_data %>%
  select(defendant, case_number, total_assessed, total_paid) %>%
  group_by(defendant) %>%
  summarise(total_paid = sum(total_paid))
```

## Calculating span of trespass arrests

```{r}             
# building for loop, leaving behind to create blank dataframe

temp4 <- trespass_data %>%
  select(case_number, defendant) %>%
  group_by(defendant, case_number) %>%
  summarise()

temp11 <- trespass_data %>%
  filter(defendant == "jerry michael kinney") %>%
  select(defendant, arrest_date) %>%
  arrange(arrest_date)
temp12 <- temp11 %>%
  slice(1) %>%
  mutate(first_arrest = arrest_date) %>%
  select(-arrest_date)
temp13 <- temp11 %>%
  slice(n()) %>%
  mutate(last_arrest = arrest_date) %>%
  select(-arrest_date)
temp14 <- temp12 %>%
  inner_join(temp13, by = "defendant")
# creating blank data frame
arrest_dates <- temp4 %>%
  filter(defendant == "nothing")

# calculating first and last arrest in filtered_crimes
for (i in names){
  temp11 <- trespass_data %>%
    filter(defendant == i) %>%
    select(defendant, arrest_date)
  temp12 <- temp11 %>%
    arrange(arrest_date) %>%
    slice(1) %>%
    mutate(first_arrest = arrest_date) %>%
    select(-arrest_date)
  temp13 <- temp11 %>%
    arrange(desc(arrest_date)) %>%
    slice(1) %>%
    mutate(last_arrest = arrest_date) %>%
    select(-arrest_date)
  temp14 <- temp12 %>%
    inner_join(temp13, by = "defendant")
  arrest_dates <- arrest_dates %>%
    bind_rows(temp14)
}

# calculating time_span
charge_dates <- arrest_dates %>%
  mutate(time_span = (last_arrest - first_arrest),
         first_charge = first_arrest,
         last_charge = last_arrest
         ) %>%
  select(first_charge, last_charge, time_span)

```

# Calculating total fines

```{r}

#total arrests and charges
temp4 <- homeless_data %>%
  select(case_number, defendant, total_assessed) %>%
  unique() %>%
  group_by(defendant) %>%
  summarise(total_assessed = sum(total_assessed))
total_charges <- homeless_data %>%
  select(case_number, defendant, arrest_charge_count) %>%
  group_by(defendant) %>%
  summarise(total_charges = n()) %>%
  inner_join(temp4)



```

# Calculating jail stays -- Broken, outdated code

```{r}

tres_jail <- trespass_data %>%
  inner_join(jail_stays, by = c("defendant" = "name", "arrest_week" = "arrest_week", "arrest_year" = "arrest_year")) %>%
  unique()
tres_jail$days_in_jail[is.na(tres_jail$days_in_jail)] <- 0

tres_jail_stays <- tres_jail %>%
  group_by(name) %>%
  summarise(tres_days_in_jail = sum(days_in_jail))

# calculating lenghth of jail stays
temp23 <- joined_jail_stays %>%
  group_by(defendant) %>%
  summarise(total_days = sum(days_in_jail))
        
temp24 <- joined_jail_stays %>%
  group_by(defendant) %>%
  summarise(average_days = mean(days_in_jail))

temp25 <- joined_jail_stays %>%
  group_by(defendant) %>%
  summarise(median_days = median(days_in_jail))

jail_stays_per_person <- temp23 %>%
  left_join(temp24) %>%
  left_join(temp25)
```

# Joining everything together

```{r}

# joining everything together
impacts <- trespass_impacts %>%
  left_join(other_crimes, by = "defendant") %>%
  left_join(other_crimes_charges, by = "defendant") %>%
  #left_join(total_cases, by = "defendant") %>%
  left_join(total_charges, by = "defendant") %>%
  left_join(other_crim_fines, by = "defendant") %>%
  left_join(total_fines, by = "defendant") %>%
  left_join(other_crim_paid, by = "defendant") %>%
  # left_join(total_paid, by = "defendant") %>%
  left_join(arrest_dates, by = "defendant") %>%
  left_join(demographics, by = "defendant") %>%
  left_join(tres_missed_court, by = "defendant") 
  #left_join(jail_stays_per_person, by = "defendant") %>%
  
#temporary <- impacts %>%  
  #select(defendant, race, gender, age, tres_charges, tres_fees, tres_paid, tres_missed_court, other_crimes_charges, other_crim_fines, total_fines, first_arrest, last_arrest, time_span) 
         
         #other_crimes, tres_missed_court, other_crimes, other_crimes_charges, other_crim_fines,  first_arrest, last_arrest, time_span, other_cases, total_fines, total_paid)
impacts$tres_missed_court[is.na(impacts$tres_missed_court)] <- 0
 #total_days, average_days, median_days, 

#impacts$total_days[is.na(impacts$total_days)] <- 0
#impacts$average_days[is.na(impacts$average_days)] <- 0
#impacts$median_days[is.na(impacts$median_days)] <- 0

# rm(temp, other_crimes, other_crimes_charges, temp4, total_cases, total_charges, other_crim_fines, total_fines, other_crim_paid, total_paid, temp11, temp12, temp13, temp14, temp15, temp16, temp17, demographics, temp19, temp20, temp21, temp22, i, arrest_dates)


```

# Kenneth Ivan Shultz III
The code that creates everything we know about Shultz's arrest.

```{r}
kenny_jail_stays <- read.csv('./data/kenny-jail-stays.csv')

kenny_jail_stays <- kenny_jail_stays %>%
    mutate(booking_date = mdy(booking_date),
         release_date = mdy(release_date),
         days_in_jail = replace_na((release_date - booking_date), 0),
         name = paste0(first, " ", last),
         arrest_week = week(booking_date),
         year = year(booking_date))

kenny_days_by_year <- kenny_jail_stays %>%
  group_by(year) %>%
  summarise(days = sum(days_in_jail))

#adjusting cost of okaloosa jail for inflation
cpi <- bls_api("CUSR0000SA0")
inflation <- inflation_adjust(2019)
inflation <- inflation %>%
  select(year, adj_value) %>%
  slice(65:74)
inflation <- inflation %>%
  mutate(jail_cost = (adj_value * (49.34)),
         year = as.double(year))

kenny_jail_cost <- kenny_days_by_year %>%
  inner_join(inflation, by = 'year')

kenny_jail_cost <- kenny_jail_cost %>%
  mutate(cost_per_year = days*jail_cost)

write.csv(kenny_jail_cost, file='./output/kenny-jail-costs.csv')


kenny <- read_csv("data/test/schultz.csv") 

temp2 <- kenny %>%
  mutate(booking_date = mdy(booking_date),
         release_date = mdy(release_date),
         days_in_jail = (release_date - booking_date)
         )

kenny_all_charges <- master_data %>%
  filter(defendant == "kenneth ivan shultz iii") %>%
  arrange(offense_date)

kenny_all_unique <- kenny_all_charges %>%
  select(case_number, defendant, total_assessed) %>%
  unique() %>%
  group_by(defendant) %>%
  summarise(sum(total_assessed))


temp4 <- temp2 %>%
  inner_join(temp3, by = c("booking_date" = "arrest_date"))

sum(temp2$days_in_jail, na.rm = TRUE)
sum(temp4$days_in_jail, na.rm = TRUE)

temp6 <- master_data %>%
  filter(defendant == "kenneth ivan shultz iii") %>%
  select(case_number, arrest_date, arrest_statute)

kenny <- temp2 %>%
  anti_join(temp3, by = c("booking_date" = "arrest_date")) %>%
  inner_join(temp6, by = c("booking_date" = "arrest_date"))
  

most_charges <- master_data %>%
  group_by(defendant) %>%
  count()




```


# Prosecutors

```{r}
# dataframes investigating prosecutors
temp1 <- trespass_data %>%
  group_by(prosecutor) %>%
  summarise(fees = sum(total_assessed))

temp2 <- trespass_data %>%
  group_by(prosecutor) %>%
  summarise(charges = n())

prosecutors <- temp1 %>%
  inner_join(temp2)

madden <- trespass_data %>%
  filter(prosecutor == "ginger bowden madden") %>%
  group_by(court_action) %>%
  count()

tres_court_action <- trespass_data %>%
  group_by(court_action) %>%
  count()

madden_cases <- trespass_data %>%
  filter(prosecutor == "ginger bowden madden")

```


# Examining which LEO agencies and prosecutors charge more

We cannot sort criminalization by city for the cities where the Okaloosa Sheriff's Office is hired to act as the law enforcement organization in cities like Destin because it appears the sheriff's office rarely inputs the city in the jurisdiction when they make arrests there.

Given the size of the sheriff's office's jurisdiction, which includes other cities in addition to Destin, I would expect the sheriff's office to be charging homeless people with these crimes more than Fort Walton Beach if Destin is criminalizing homelessness more than Fort Walton Beach.

```{r}

# comparing the four rates of arrests among LEO agencies in Okaloosa
temp <- master_data %>%
  group_by(arresting_agency) %>%
  summarise(total_count = n()) %>%
  mutate(temp = (total_count/(nrow(master_data)))*100) %>%
  mutate(perc_of_total = round(temp, 2)) %>%
  select(-temp)

temp2 <- homeless_data %>%
  group_by(arresting_agency) %>%
  summarise(homeless_count = n()) %>%
  mutate(temp = (homeless_count/(nrow(homeless_data))*100)) %>% 
  mutate(perc_of_homeless = round(temp, 2)) %>%
  select(-temp)

temp3 <- trespass_data %>%
  group_by(arresting_agency) %>%
  summarise(trespass_count = n()) %>%
  mutate(temp = (trespass_count/(nrow(trespass_data))*100)) %>% 
  mutate(perc_of_trespass = round(temp, 2)) %>%
  select(-temp)

temp4 <- filtered_crimes %>%
  group_by(arresting_agency) %>%
  summarise(crimes_count = n()) %>%
  mutate(temp = (crimes_count/(nrow(filtered_crimes))*100)) %>% 
  mutate(perc_of_filtered = round(temp, 2)) %>%
  select(-temp)

agency <- temp %>%
  inner_join(temp2, by = "arresting_agency") %>%
  inner_join(temp3, by = "arresting_agency") %>%
  inner_join(temp4, by = "arresting_agency") %>%
  select(arresting_agency, perc_of_total, perc_of_homeless, perc_of_trespass, perc_of_filtered) %>%
  arrange(desc(perc_of_filtered))

rm(temp, temp2, temp3)

```

```{r}

# comparing who charged the most filtered crimes

temp <- filtered_crimes %>%
  filter(str_detect(arresting_agency, "oka")) %>%
  group_by(arrest_charge) %>%
  summarise(sheriff_charges = n())

temp2 <- filtered_crimes %>%
  filter(str_detect(arresting_agency, "ft")) %>%
  group_by(arrest_charge) %>%
  summarise(fwb_charges = n())

temp3 <- filtered_crimes %>%
  filter(str_detect(arresting_agency, "nice")) %>%
  group_by(arrest_charge) %>%
  summarise(nice_charges = n())

temp4 <- filtered_crimes %>%
  filter(str_detect(arresting_agency, "crest")) %>%
  group_by(arrest_charge) %>%
  summarise(crest_charges = n())

temp5 <- filtered_crimes %>%
  filter(str_detect(arresting_agency, "valp")) %>%
  group_by(arrest_charge) %>%
  summarise(valp_charges = n())
  
agency_filtered <- temp %>%
  full_join(temp2, by = "arrest_charge") %>%
  full_join(temp3, by = "arrest_charge") %>%
  full_join(temp4, by = "arrest_charge") %>%
  full_join(temp5, by = "arrest_charge") 
# transforming NAs to zeroes
agency_filtered[is.na(agency_filtered)] <- 0
#finishing math  
agency_filtered <- agency_filtered %>%
  mutate(temp = (fwb_charges + sheriff_charges + nice_charges + crest_charges + valp_charges)) %>%
  mutate(perc_fwb = (fwb_charges/temp)*100,
         statute = arrest_charge) %>%
  arrange(desc(perc_fwb)) %>%
  select(statute, nice_charges, crest_charges, valp_charges, sheriff_charges, fwb_charges, perc_fwb) 

rm(temp, temp2, temp3, temp4, temp5)

```

```{r}

# comparing the rates of prosecutors on homeless
temp <- master_data %>%
  group_by(prosecutor) %>%
  summarise(total_count = n()) %>%
  mutate(temp = (total_count/(nrow(master_data))*100)) %>%
  mutate(perc_of_total = round(temp, 2)) %>%
  select(-temp)

temp2 <- homeless_data %>%
  group_by(prosecutor) %>%
  summarise(homeless_count = n()) %>%
  mutate(temp = (homeless_count/(nrow(homeless_data))*100)) %>% 
  mutate(perc_of_homeless = round(temp, 2)) %>%
  select(-temp)

temp3 <- crimes_data %>%
  group_by(prosecutor) %>%
  summarise(crimes_count = n()) %>%
  mutate(temp = (crimes_count/(nrow(crimes_data))*100)) %>% 
  mutate(perc_of_crimes = round(temp, 2)) %>%
  select(-temp)

temp4 <- filtered_crimes %>%
  group_by(prosecutor) %>%
  summarise(crimes_count = n()) %>%
  mutate(temp = (crimes_count/(nrow(filtered_crimes))*100)) %>% 
  mutate(perc_of_filtered = round(temp, 2)) %>%
  select(-temp)

prosecutor <- temp %>%
  inner_join(temp2, by = "prosecutor") %>%
  inner_join(temp3, by = "prosecutor") %>%
  inner_join(temp4, by = "prosecutor") %>%
  select(prosecutor, perc_of_total, perc_of_homeless, perc_of_crimes, perc_of_filtered) %>%
  arrange(desc(perc_of_filtered))

rm(temp)
rm(temp2)
rm(temp3)

```

```{r}

# comparing the rates of prosecutors on homeless
temp <- master_data %>%
  group_by(judge) %>%
  summarise(total_count = n()) %>%
  mutate(temp = (total_count/(nrow(master_data))*100)) %>%
  mutate(perc_of_total = round(temp, 2)) %>%
  select(-temp)

temp2 <- homeless_data %>%
  group_by(judge) %>%
  summarise(homeless_count = n()) %>%
  mutate(temp = (homeless_count/(nrow(homeless_data))*100)) %>% 
  mutate(perc_of_homeless = round(temp, 2)) %>%
  select(-temp)

temp3 <- crimes_data %>%
  group_by(judge) %>%
  summarise(crimes_count = n()) %>%
  mutate(temp = (crimes_count/(nrow(crimes_data))*100)) %>% 
  mutate(perc_of_crimes = round(temp, 2)) %>%
  select(-temp)

temp4 <- filtered_crimes %>%
  group_by(judge) %>%
  summarise(crimes_count = n()) %>%
  mutate(temp = (crimes_count/(nrow(filtered_crimes))*100)) %>% 
  mutate(perc_of_filtered = round(temp, 2)) %>%
  select(-temp)

temp5 <-  filtered_crimes %>%
  group_by(judge) %>%
  summarise(homeless_levied = sum(total_assessed))


judge <- temp %>%
  inner_join(temp2, by = "judge") %>%
  inner_join(temp3, by = "judge") %>%
  inner_join(temp4, by = "judge") %>%
  inner_join(temp5, by = "judge") %>%
  select(judge, perc_of_total, perc_of_homeless, perc_of_crimes, perc_of_filtered, homeless_levied, crimes_count) %>%
  arrange(desc(perc_of_filtered))

rm(temp)
rm(temp2)
rm(temp3)

```



```{r}
# Only specific crimes
# c("316.130", "316.2045", "810.08", "810.09", "856.011"))

temp <- crimes_data %>%
  filter(str_detect(arresting_agency, "oka")) %>%
  group_by(arrest_charge) %>%
  summarise(sheriff_charges = n())

temp2 <- crimes_data %>%
  filter(str_detect(arresting_agency, "ft")) %>%
  group_by(arrest_charge) %>%
  summarise(fwb_charges = n())
  
temp %>%
  inner_join(temp2, by = "arrest_charge") %>%
  select(arrest_charge, fwb_charges, sheriff_charges) %>%
  mutate(temp = (fwb_charges + sheriff_charges)) %>%
  mutate(perc_sheriff = (sheriff_charges/temp)*100) %>%
  arrange(desc(perc_sheriff))


rm(temp)
rm(temp2)


```

```{r}

affray <- master_data %>%
  filter(str_detect(arrest_charge, "870.01"))


```

# Time
```{r}

temp <- filtered_crimes %>%
  group_by(offense_year) %>%
  summarize(total_charges = n())

temp2 <- filtered_crimes %>%
  filter(str_detect(arresting_agency, "oka")) %>%
  group_by(offense_year) %>%
  summarize(shr_charges = n())

temp3 <- filtered_crimes %>%
  filter(str_detect(arresting_agency, "ft")) %>%
  group_by(offense_year) %>%
  summarize(ftw_charges = n())

temp4 <- filtered_crimes %>%
  filter(str_detect(arresting_agency, "nice")) %>%
  group_by(offense_year) %>%
  summarize(nice_charges = n())

temp5 <- filtered_crimes %>%
  filter(str_detect(arresting_agency, "crest")) %>%
  group_by(offense_year) %>%
  summarize(crest_charges = n())

temp6 <- filtered_crimes %>%
  filter(str_detect(arresting_agency, "valp")) %>%
  group_by(offense_year) %>%
  summarize(valp_charges = n())

time <- temp %>%
  full_join(temp2, by = "offense_year") %>%
  full_join(temp3, by = "offense_year") %>%
  full_join(temp4, by = "offense_year") %>%
  full_join(temp5, by = "offense_year") %>%
  full_join(temp6, by = "offense_year")

rm(temp, temp2, temp3, temp4, temp5, temp6)




```

```{r}
temp <- master_data %>%
  group_by(offense_year) %>%
  summarize(total_charges = n())

temp2 <- master_data %>%
  filter(str_detect(arresting_agency, "oka")) %>%
  group_by(offense_year) %>%
  summarize(shr_charges = n())

temp3 <- master_data %>%
  filter(str_detect(arresting_agency, "ft")) %>%
  group_by(offense_year) %>%
  summarize(ftw_charges = n())

temp4 <- master_data %>%
  filter(str_detect(arresting_agency, "nice")) %>%
  group_by(offense_year) %>%
  summarize(nice_charges = n())

temp5 <- master_data %>%
  filter(str_detect(arresting_agency, "crest")) %>%
  group_by(offense_year) %>%
  summarize(crest_charges = n())

temp6 <- master_data %>%
  filter(str_detect(arresting_agency, "valp")) %>%
  group_by(offense_year) %>%
  summarize(valp_charges = n())

all_cr_time <- temp %>%
  full_join(temp2, by = "offense_year") %>%
  full_join(temp3, by = "offense_year") %>%
  full_join(temp4, by = "offense_year") %>%
  full_join(temp5, by = "offense_year") %>%
  full_join(temp6, by = "offense_year")

rm(temp, temp2, temp3, temp4, temp5, temp6)



```


```{r}

recent_filtered <- filtered_crimes %>%
  filter(offense_year == c("2018", "2019"))

temp <- recent_filtered %>%
  group_by(arrest_month) %>%
  summarise(total_charges = n())

temp2 <- recent_filtered %>%
  filter(str_detect(arresting_agency, "oka")) %>%
  group_by(arrest_month) %>%
  summarize(shr_charges = n())

temp3 <- recent_filtered %>%
  filter(str_detect(arresting_agency, "ft")) %>%
  group_by(arrest_month) %>%
  summarize(ftw_charges = n())

temp4 <- recent_filtered %>%
  filter(str_detect(arresting_agency, "nice")) %>%
  group_by(arrest_month) %>%
  summarize(nice_charges = n())

temp5 <- recent_filtered %>%
  filter(str_detect(arresting_agency, "crest")) %>%
  group_by(arrest_month) %>%
  summarize(crest_charges = n())

temp6 <- recent_filtered %>%
  filter(str_detect(arresting_agency, "valp")) %>%
  group_by(arrest_month) %>%
  summarize(valp_charges = n())

recent_month <- temp %>%
  full_join(temp2, by = "arrest_month") %>%
  full_join(temp3, by = "arrest_month") %>%
  full_join(temp4, by = "arrest_month") %>%
  full_join(temp5, by = "arrest_month") %>%
  full_join(temp6, by = "arrest_month")

recent_month[is.na(recent_month)] <- 0

rm(temp, temp2, temp3, temp4, temp5, temp6)
```

```{r}

temp <- recent_filtered %>%
  filter(str_detect(arresting_agency, "oka")) 

recent_breakdown <- temp %>%
  group_by(city) %>%
  summarise(count = n())

```

# Second look at all tresspassing laws charged in Okaloosa

```{r}

homeless_tresspass <- homeless_data %>%
  filter(str_detect(arrest_charge, "810")) %>%
  group_by(arrest_charge, arrest_statute) %>%
  summarise(count = n()) %>%
  arrange(desc(count))

tresspass <- crimes_data %>%
  filter(str_detect(arrest_charge, "810")) %>%
  group_by(arrest_charge, arrest_statute) %>%
  summarise(count = n()) %>%
  arrange(desc(count))

```


```{r}

tres_over_time <- trespass_data %>%
  group_by(arrest_year) %>%
  summarise(charges = n())

```


```{r}
test <- master_data %>%
  filter(str_detect(arrest_statute,"TRES")) %>%
  filter(!str_detect(arrest_charge,"2c|2d|97|13")) 
%>%
  group_by(arrest_year) %>%
  summarise(count=n())



```

```{r}

trespass_data %>%
  group_by(arrest_year) %>%
  count()
```



```{r}



#joining on booking_date and arrest_date
exact_match <- jail_stays %>%
  unique() %>% 
  select(-address) %>%
  inner_join(master_data, by = c("booking_date" = "arrest_date", "name" = "defendant"))

#joining on next_day and arrest_date
next_day <- jail_stays %>%
  unique() %>% 
  select(-address) %>%
  inner_join(master_data, by = c("booking_date" = "next_day", "name" = "defendant"))

#joining on arrest_week
arrest_week <- jail_stays %>%
  inner_join(master_data, by = c("arrest_week" = "arrest_week", "name" = "defendant"))

no_match <- jail_stays %>%
  anti_join(master_data, by = c("arrest_week" = "arrest_week", "name" = "defendant"))




  
```


```{r}

days_kenny <- kenny %>%
  mutate(days = (release_date - booking_date))


```


```{r}

case_numbers <- kenny_all_charges %>%
  select(case_number) %>%
  unique()

write.csv(case_numbers, file = 'output/kenny-case-numbers.csv')

```


# Saving

```{r}

# This loop creates files with a list of all charges filed in the database
for (i in names){
  filename <- paste0("individuals/", i, ".csv")
  temp <- master_data %>%
    filter(defendant == i) %>%
    group_by(arrest_statute, arrest_charge) %>%
    summarise(count = n()) %>%
    arrange(desc(count))
  write.csv(temp, file = filename)
}
```

```{r}

# homeless people impacted
write.csv(impacts, file = "output/impacts.csv")
# what is getting charged
write.csv(homeless_charges, file = "output/homeless_charges.csv")
# who charges the most per criminalization statute?
write.csv(agency_filtered, file = "output/agency-filtered.csv")
# who is charging it
write.csv(agency, file = "output/agency.csv")
# who is prosecuting these people?
write.csv(prosecutor, file = "output/prosecutor.csv")
write.csv(prosecutors, file = "output/tres-charges-by-prosecutors.csv" )
# who is sentencing these people?
write.csv(judge, file = "output/judge.csv")
# yearly trend?
write.csv(time, file = "output/time.csv")
write.csv(recent_month, file = "output/recent_month.csv")
write.csv(recent_breakdown, file = "output/recent_breakdown.csv")
write.csv(kenny_all_charges, file = "output/kenny_all_charges.csv")
write.csv(kenny_all_unique, file = "output/kenny_all_unique_costs.csv")
write.csv(kenny_jail_stays, file = "output/kenny_jail_stays.csv")
write.csv(trespass_impacts, file = "output/trespass_impacts.csv")
# what is the race breakdown of the people?
write.csv(tresspass, file = "output/trespass.csv")

write.csv(names, file = "names.csv")

write.csv(master_data, file = "master-data.csv")

```